Population structure and phenotypic variation of \emph{Sclerotinia sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United States

### 2017-11-28 doc/manuscript/paper.tex
add citations

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017) and are openly available and reproducible at
\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
microsatellite loci genotyped, five included compound repeats, which
made it challenging to accurately/confidently bin alleles into fragment
sizes expected for each locus based on the described repeat motif. Loci
with compound repeats were removed for the reported statistics. To
ensure the integrity of the results we additionally processed these loci
and included them in concurrent analyses. We assessed the power of our
11 markers by generating a genotype accumulation curve in the R package
\emph{poppr} version 2.5.0, looking for evidence of saturation, which
would indicate that loci were sufficiently sampled to adequately
represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
et al., 2015). To avoid including isolates potentially collected from
the same plant (which increases the probability of collecting sclerotia
from the same point of infection more than once), data were
clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
duplicated genotypes were reduced to a single observation when they were
collected in the same year from the same host cultivar located in the
same source field (wmn or producer)---for subsequent analysis. We
assessed haplotype diversity by calculating Stoddart and Taylor's index
(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
(\(E_5\)), and the expected number of multilocus haplotypes
(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
et al., 2003). To assess the potential for random mating, we tested for
linkage disequilibrium with the index of association, \(I_A\) and its
standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
diversity and linkage disequilibrium were calculated in \emph{poppr}
(Kamvar et al., 2014).

@@

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/Fig1.pdf}
\caption{Associations between Mycelial Compatibility Groups and
Multilocus Haplotypes. \textbf{A)} Barplot of Mycelial Compatibility
Group (MCG) abundance in descending order. Singletons (46) were
truncated, leaving 41 MCGs. White bars represent sample counts and grey
bars represent counts of unique multilocus haplotypes (MLH). The
transparency of the bars represent the evenness of the distribution of
the MLHs within a given MCG. A dashed box surrounds the eight most
common MCGs representing \textgreater{}~51\% of the data. \textbf{B)}
Full graph-representation of the relationship between MCGs (open
circles) and MLHs (filled circles). Details in Fig. \ref{fullgraph}.
\textbf{C)} A subset of \textbf{B} representing the 8 most common MCGs
and their associated MLHs (dashed box in \textbf{A}). Filled nodes
(circles) represent MLHs and open nodes represent MCGs. Node area scaled
to the number of samples represented (range: 1--73). Numbers inside
nodes are the MLH/MCG label (if n \textgreater{} 1). Edges (arrows)
point from MLH to MCG where the weight (thickness) of the edge
represents the number of shared isolates (range: 1--19). Edges extending
from MLHs displayed to other MCGs are not shown.}\label{biggraph}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig2.pdf}
\caption{Biplot showing five most influential explanatory variables
(arrows) overlayed on the first two eigenvectors of distance based
redundancy analysis of \emph{Sclerotinia sclerotiorum} isolates. The
length of the arrows are directly proportional to the strength of the
correlation between explanatory and molecular variables. Open circles
represent the 318 clone-corrected haplotypes in ordination
space.}\label{rda-biplot}
\end{figure}

\begin{figure}
\centering
\includegraphics{../../results/figures/publication/Fig3.pdf}
\caption{Strip plot of aggressiveness by population arranged in
descending order of mean aggressiveness for all populations with N
\textgreater{} 5. White bars represent mean value. Circles represent
individual isolates where filled circles are isolates from white mold
screening nurseries (wmn) and open circles are isolates from producer
fields.}\label{aggressiveness-boxplot}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig4.pdf}
\caption{Network of populations (nodes/circles) and their shared
multilocus haplotypes (MLH) (edges/lines) genotyped over 11 loci. Each
node is labeled with \textbf{name (number of MLHs shared/number of MLHs
total).} The shade and area of the nodes are proportional to the number
of unique MLHs within the node and the inner nodes are proportional to
the number of private MLHs to the region (bottom legend). Each edge
represents a single MLH where its thickness represents the number of
populations that share the MLH and the shade represents the value of
\(P_{sex}\), or the probability of encountering that MLH from two
independent meiotic events.}\label{community-graph}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/Fig5.pdf}
\caption{Discriminant Analysis of Principal Components (DAPC) on regions
showing that Mexico is differentiated from other populations.
\textbf{A)} Scatter plot of first two components from DAPC. Points
represent observed individuals connected to the population centroids
with ellipses representing a 66\% confidence interval for a normal
distribution. The center of each component is represented as black grid
lines. \textbf{B)} Mean population assignment probability from the DAPC
for all populations with N \textgreater{} 10 (facets). Populations
represented along the horizontal axis and probability of assignment on
the vertical. Numbers next to source populations indicate population
size. All values sum to one.}\label{DAPC}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig6.pdf}
\caption{Scatter plot of Discriminant Analysis of Principal Components
(DAPC) on Regions and Years showing non-differentiated temporal
variation NE and MI and temporal variation in WA and CA. Points (text
labels) represent observed individuals connected to the population
centroids with ellipses representing a 66\% confidence interval for a
normal distribution. The center of each component is represented as
black grid lines. A more detailed view is shown in Fig.
\ref{DAPC-RY-FULL}.}\label{DAPC-RY}
\end{figure}

@@

@@

>>>>

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017). These analyses were rendered as dynamic
documents with the R packages \emph{knitr} (version 1.17) and
\emph{ezknitr} (version 0.6) and are openly available and reproducible
at \url{https://github.com/everhartlab/sclerotinia-366/} (Attali, 2016;
Xie, 2017). Of the 16 microsatellite loci genotyped, five included
compound repeats, which made it challenging to accurately/confidently
bin alleles into fragment sizes expected for each locus based on the
described repeat motif. Loci with compound repeats were removed for the
reported statistics. To ensure the integrity of the results we
additionally processed these loci and included them in concurrent
analyses. We assessed the power of our 11 markers by generating a
genotype accumulation curve in the R package \emph{poppr} version 2.5.0,
looking for evidence of saturation, which would indicate that loci were
sufficiently sampled to adequately represent the full set of haplotypes
(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). To avoid including
isolates potentially collected from the same plant (which increases the
probability of collecting sclerotia from the same point of infection
more than once), data were clone-corrected on a hierarchy of
Region/Source/Host/Year---meaning that duplicated genotypes were reduced
to a single observation when they were collected in the same year from
the same host cultivar located in the same source field (wmn or
producer)---for subsequent analysis. We assessed haplotype diversity by
calculating Stoddart and Taylor's index (\emph{G}) (Stoddart \& Taylor,
1988), Shannon's index (\emph{H}) (Shannon, 1948), Simpson's index
(\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
number of multilocus haplotypes (\emph{eMLH}) (Hurlbert, 1971; Heck et
al., 1975; Pielou, 1975; Grünwald et al., 2003). To assess the potential
for random mating, we tested for linkage disequilibrium with the index
of association, \(I_A\) and its standardized version, \(\bar{r}_d\)
using 999 permutations (Brown et al., 1980; Smith et al., 1993; Agapow
\& Burt, 2001). Both haplotype diversity and linkage disequilibrium were
calculated in \emph{poppr} (Kamvar et al., 2014).

@@

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mcg-subgraph-with-context.pdf}
\caption{Associations between Mycelial Compatibility Groups and
Multilocus Haplotypes. \textbf{A)} Barplot of Mycelial Compatibility
Group (MCG) abundance in descending order. Singletons (46) were
truncated, leaving 41 MCGs. White bars represent sample counts and grey
bars represent counts of unique multilocus haplotypes (MLH). The
transparency of the bars represent the evenness of the distribution of
the MLHs within a given MCG. A dashed box surrounds the eight most
common MCGs representing \textgreater{}~51\% of the data. \textbf{B)}
Full graph-representation of the relationship between MCGs (open
circles) and MLHs (filled circles). Details in Fig. \ref{fullgraph}.
\textbf{C)} A subset of \textbf{B} representing the 8 most common MCGs
and their associated MLHs (dashed box in \textbf{A}). Filled nodes
(circles) represent MLHs and open nodes represent MCGs. Node area scaled
to the number of samples represented (range: 1--73). Numbers inside
nodes are the MLH/MCG label (if n \textgreater{} 1). Edges (arrows)
point from MLH to MCG where the weight (thickness) of the edge
represents the number of shared isolates (range: 1--19). Edges extending
from MLHs displayed to other MCGs are not shown.}\label{biggraph}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Figure7Z.pdf}
\caption{Biplot showing five most influential explanatory variables
(arrows) overlayed on the first two eigenvectors of distance based
redundancy analysis of \emph{Sclerotinia sclerotiorum} isolates. The
length of the arrows are directly proportional to the strength of the
correlation between explanatory and molecular variables. Open circles
represent the 318 clone-corrected haplotypes in ordination
space.}\label{rda-biplot}
\end{figure}

\begin{figure}
\centering
\includegraphics{../../results/figures/publication/aggressiveness.pdf}
\caption{Strip plot of aggressiveness by population arranged in
descending order of mean aggressiveness for all populations with N
\textgreater{} 5. White bars represent mean value. Circles represent
individual isolates where filled circles are isolates from white mold
screening nurseries (wmn) and open circles are isolates from producer
fields.}\label{aggressiveness-boxplot}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/mlg-11.pdf}
\caption{Network of populations (nodes/circles) and their shared
multilocus haplotypes (MLH) (edges/lines) genotyped over 11 loci. Each
node is labeled with \textbf{name (number of MLHs shared/number of MLHs
total).} The shade and area of the nodes are proportional to the number
of unique MLHs within the node and the inner nodes are proportional to
the number of private MLHs to the region (bottom legend). Each edge
represents a single MLH where its thickness represents the number of
populations that share the MLH and the shade represents the value of
\(P_{sex}\), or the probability of encountering that MLH from two
independent meiotic events.}\label{community-graph}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/DAPC.pdf}
\caption{Discriminant Analysis of Principal Components (DAPC) on regions
showing that Mexico is differentiated from other populations.
\textbf{A)} Scatter plot of first two components from DAPC. Points
represent observed individuals connected to the population centroids
with ellipses representing a 66\% confidence interval for a normal
distribution. The center of each component is represented as black grid
lines. \textbf{B)} Mean population assignment probability from the DAPC
for all populations with N \textgreater{} 10 (facets). Populations
represented along the horizontal axis and probability of assignment on
the vertical. Numbers next to source populations indicate population
size. All values sum to one.}\label{DAPC}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/dapc_region_year_micanewa.pdf}
\caption{Scatter plot of Discriminant Analysis of Principal Components
(DAPC) on Regions and Years showing non-differentiated temporal
variation NE and MI and temporal variation in WA and CA. Points (text
labels) represent observed individuals connected to the population
centroids with ellipses representing a 66\% confidence interval for a
normal distribution. The center of each component is represented as
black grid lines. A more detailed view is shown in Fig.
\ref{DAPC-RY-FULL}.}\label{DAPC-RY}
\end{figure}

@@

\hypertarget{ref-attali2016ezknitr}{}
Attali, D. (2016). ezknitr: Avoid the typical working directory pain
when using ``knitr''. \emph{The Journal of Open Source Software},
\emph{1}(5), 75. \url{https://doi.org/10.21105/joss.00075}

@@

\hypertarget{ref-xie2017knitr}{}
Xie, Y. (2017). \emph{knitr: A general-purpose package for dynamic
report generation in R}. Retrieved from \url{https://yihui.name/knitr/}
